---
title: "Lights, Camera, Action! - Predicting Gross Revenue of Movies"
author: "Rated R: Carlie Scheer, Christina Lee, Jerry Lin, Samir Travers"
date: "10/27/2024"
format: pdf
execute: 
  warning: false
  message: false
  echo: false
editor: visual
---

```{r}
#| label: load-pkg-data
library(tidyverse)
library(tidymodels)
library(knitr)
library(readr)
library(GGally)
library(dplyr)

movies <- read_csv("data/final_vote_filtered_movie_data.csv")
```

## Introduction

Movies have been at the heart of pop culture for nearly a century. And while movies have changed a lot over the years, varying in length, theme, due to technological advances, and more, one thing has never changed–movies have remained as a dominant part of the entertainment industry. Yet, ever since the COVID-19 pandemic, the North American box office for movies has not recovered–it seems that people just aren’t going to movies as much as they used to. As mentioned in [@boxofficemojo2024], domestic box office gross revenue in 2019 reached nearly \$11.4 billion, while the yearly total from 2023 topped off at \$8.9 billion after a very slow two years in 2020 and 2021. With moviegoers back on the rise, the film industry and box offices are hopeful to end the decade on a high note and return back to their original state. As such, we are interested in investigating the trends in gross revenue for top movies and how various elements impact the amount of money that a movie makes.

Research Question: What are the key factors and metrics that most significantly influence a movie's gross revenue?

Motivation: As a group of movie enthusiasts, movie popularity has always been a large source of curiosity. We’ve seen terrible movies whose popularity we find puzzling, and unknown masterpieces that do not get the attention they deserve. What is it that separates the two? Is it primarily the “goodness” of the movie that determines the money it brings in, or are there other factors that are equally influential? Is the 1.4 billion dollar success of Alvin and the Chipmunks (an objectively pretty bad movie) just an outlier, or reflective of a deep and complex relationship between a movie’s characteristics and its gross profit? Through our analysis, we hope to get to the bottom of these questions, and gain a deeper understanding of what truly drives a movie's success or failure in the box office.

Hypotheses: We hypothesize movies that have high popularity scores and ratings, substantial budgets, and broad audience appeal will be more likely to generate higher gross revenue. We also hypothesize movies with more than one genre will be more likely to generate higher gross revenue. Movies with a runtime between 1.5-2 hours will be more likely to generate higher gross revenue.

## Data description

The source of the dataset is taken from TMDb (The Movie Database), which is an online, public database for movies that contains information such as ratings, runtime, cast, director, and box office performance. We discovered this dataset on Kaggle, and it is updated daily by user asaniczka, who takes the data from TMDb [@asaniczka2023].

There were 953629 observations in the original dataset, each representing a single film. There were 24 columns in this dataset, with some of the most important being the name (an identifier), release year, runtime, genres, rating, votes, and revenue.

## Exploratory data analysis

We filtered the original dataset outside of this project first because the original dataset of almost 1 million rows was far too large to be able to run efficiently in RStudio; as such, we filtered for where the movies had more than 100 votes, meaning more than 100 people submitted ratings for the movie. After filtering, there are now 18086 observations, and we were able to import this filtered dataset into Rstudio (The file name was "final_vote_filtered_movie_data.csv").

We are going to transform the variables spoken_languages and genres. For spoken_languages, we want to make a categorical variable based on whether more than one language is spoken in the movie or not (True = more than one language, False = just one language). For genres, we want to take a similar approach to see whether the movie is tagged for more than one genre (True = more than one genre, False = just one genre). Additionally, we want to transform the month, date, year format of the release date to just be the year. Lastly, before we begin EDA and further analysis, we are going to remove all 0s in the revenue and budget columns because we are not able to accurately impute the values; we have evidence to suggest that the original author of the dataset just put 0s instead of NAs. 

```{r}
#Question to ask: when copying from proposal to written report qmd, do we need everything from the proposal or only the intro section? 

#New column where TRUE for multiple languages, FALSE for 1 language
movies <- movies |>
  mutate(multi_language = if_else(str_detect(spoken_languages, ','), TRUE, FALSE))

#New column where TRUE for multiple genres, FALSE for 1 genre
movies <- movies |>
  mutate(multi_genre = if_else(str_detect(genres, ','), TRUE, FALSE))

#New column for year of the movie release
movies <- movies |>
  mutate(year = as.numeric(format(as.Date(release_date), "%Y")))

#Filter down df for movies where revenue and budget is not 0
movies <- movies |>
  filter(revenue > 0 & budget > 0)

#Drop NAs for multi-languages
na_rows <- movies |> 
  filter(is.na(multi_language))
num_na_rows <- nrow(na_rows) # 4 rows with NAs
movies <- movies |> 
  drop_na(multi_language)
```

Below is basic initial EDA of our untransformed response variable, revenue.

```{r}

#Histogram
movies |>
  ggplot(aes(x = revenue)) +
  geom_histogram(bins = 20) + 
  labs(
    title = "Histogram of Revenue",
    x = "Revenue (in billions of dollars)", 
    y = "Count"
  ) +
  scale_x_continuous(labels = label_dollar(scale_cut = cut_short_scale()))

#Boxplot
movies |>
  ggplot(aes(x = revenue)) +
  geom_boxplot() + 
  labs(
    title = "Boxplot of Revenue",
    x = "Revenue (in billions of dollars)"
  ) +
  scale_x_continuous(labels = label_dollar(scale_cut = cut_short_scale()))

#Summary statistics 
movies |>
  summarise(min = min(revenue), 
            q1 = quantile(revenue, 0.25), 
            median = median(revenue), 
            q3 = quantile(revenue, 0.75), 
            max = max(revenue), 
            mean = mean(revenue), 
            sd = sd(revenue)) |>
  kable(digits = 3)


```

The distribution of revenue is strongly skewed to the right. The boxplot suggests that we have a significant amount of outliers. The median is \$18,553,131 and our interquartile range is \$48,844,861.

## Analysis approach

Variables of interest:

vote_average: This is the average rating given to the movie by viewers. We expect this to be correlated with the response variable, as good ratings will encourage more people to go see the movie, and if someone likes the movie they will be more likely to recommend it to their friends.

spoken_languages: This is the number of different languages spoken in the movie. We hypothesize that if a movie has more languages in it, it will appeal to a broader audience and therefore earn greater revenue. We created a new variable from this variable called "multi_language" that categorizes movies with multiple languges as "TRUE" and movies with one language as "FALSE."

budget: This is the amount of money spent on producing the movie. We expect movies with a larger budget to earn more revenue, as the production company has more resources to put into the movie.

runtime: This is the length of the movie. We think that people may be more willing to see shorter/average length movies, as it is less of a time commitment for them.

release_date: This indicates the day, month, and year that the movie was released. We expect that the year will greatly impact gross revenue because, for instance, movies released right before or during the pandemic will have very low gross revenue as no one was able to go to the theaters. We created a new variable from this variable called "year" that only accounts for only the year (yyyy) that the movie was released in instead of the original date (yyyy/mm/dd).

genres: This is a genre or list of genres associated with the movie. We expect movies tagged with more than one genre to appeal to a wider audience, thus driving up the number of people who go to see the movie and increasing gross revenue. We created a new variable from this variable called "multi_genre" that categorizes movies with multiple genres as "TRUE" and movies with one genre as "FALSE."


Method of approach:

We plan to use multiple linear regression to predict gross revenue, but we will log transform gross revenue because we can see from our EDA that revenue is extremely skewed in its original form.

```{r}
movies <- movies |>
  mutate(log_revenue = log(movies$revenue))
```

## Data dictionary

The data dictionary can be found here: [data/README.md](data/README.md)

## Univariate EDA

```{r}
#Histogram
movies |>
  ggplot(aes(x = log_revenue)) +
  geom_histogram(bins = 20) + 
  labs(
    title = "Histogram of Logged revenue",
    x = "Logged Revenue (in billions of dollars)", 
    y = "Count"
  ) +
  scale_x_continuous(labels = label_dollar(scale_cut = cut_short_scale()))

#Boxplot
movies |>
  ggplot(aes(x = log_revenue)) +
  geom_boxplot() + 
  labs(
    title = "Boxplot of Logged revenue Revenue",
    x = "Logged Revenue (in billions of dollars)"
  ) +
  scale_x_continuous(labels = label_dollar(scale_cut = cut_short_scale()))

#Summary statistics 
movies |>
  summarise(min = min(log_revenue), 
            q1 = quantile(log_revenue, 0.25), 
            median = median(log_revenue), 
            q3 = quantile(log_revenue, 0.75), 
            max = max(log_revenue), 
            mean = mean(log_revenue), 
            sd = sd(log_revenue),
            IQR = q3-q1)|>
  kable(digits = 3)
```

Following a log transformation, our response variable, logged revenue, has a slight left skew but is unimodal and very roughly normal. It has a significant number of outliers on the left side, and has a median of 16.74 logged billions of dollars and an IQR of 2.91.

```{r}

# Histogram
movies |>
  ggplot(aes(x = vote_average)) +
  geom_histogram(bins = 20) + 
  labs(
    title = "Histogram of Average Votes",
    x = "Average Vote (in points)", 
    y = "Count"
  ) +
  scale_x_continuous(labels = label_number())

# Boxplot
movies |>
  ggplot(aes(x = vote_average)) +
  geom_boxplot() + 
  labs(
    title = "Boxplot of Average Votes",
    x = "Average Vote (in points)"
  ) +
  scale_x_continuous(labels = label_number())

# Summary statistics 
movies |>
  summarise(
    min = min(vote_average), 
    q1 = quantile(vote_average, 0.25), 
    median = median(vote_average), 
    q3 = quantile(vote_average, 0.75), 
    max = max(vote_average), 
    mean = mean(vote_average), 
    sd = sd(vote_average),
    IQR = q3 - q1
  ) |>
  kable(digits = 3)

```

The predictor variable vote_average, measuring the average rating given to a movie by users, is roughly normally distirbuted and is unimodal. It has a few outliers on the right side, and has many on the left. The median is 6.56 points (out of ten), and the IQR is 1.09.

```{r}
# Histogram
movies |>
  ggplot(aes(x = runtime)) +
  geom_histogram(bins = 40) + 
  labs(
    title = "Histogram of Movie Runtimes",
    x = "Runtime (in minutes)", 
    y = "Count"
  ) + scale_x_continuous(limits = c(0, 250), labels = label_number())

# Boxplot
movies |>
  ggplot(aes(x = runtime)) +
  geom_boxplot() + 
  labs(
    title = "Boxplot of Movie Runtimes",
    x = "Runtime (in minutes)"
  ) +
  scale_x_continuous(labels = label_number())

# Summary statistics 
movies |>
  summarise(
    min = min(runtime), 
    q1 = quantile(runtime, 0.25), 
    median = median(runtime), 
    q3 = quantile(runtime, 0.75), 
    max = max(runtime), 
    mean = mean(runtime), 
    sd = sd(runtime),
    IQR = q3 - q1
  ) |>
  kable(digits = 3)

```

The predictor variable movie_runtime, is roughly normally distributed and unimodal, however it has a right skew. There are numerous outliers on both the left and right, including one very large outlier of 566 minutes. The median is 105 minutes, and the IQR is 23.

```{r}
# Histogram
movies |>
  ggplot(aes(x = year)) +
  geom_histogram(bins = 20) + 
  labs(
    title = "Histogram of Movie Release Years",
    x = "Release Year (in years)", 
    y = "Count"
  ) +
  scale_x_continuous(labels = label_number())

# Boxplot
movies |>
  ggplot(aes(x = year)) +
  geom_boxplot() + 
  labs(
    title = "Boxplot of Movie Release Years",
    x = "Release Year (in years)"
  ) +
  scale_x_continuous(labels = label_number())

# Summary statistics 
movies |>
  summarise(
    min = min(year), 
    q1 = quantile(year, 0.25), 
    median = median(year), 
    q3 = quantile(year, 0.75), 
    max = max(year), 
    mean = mean(year), 
    sd = sd(year),
    IQR = q3 - q1
  ) |>
  kable(digits = 3)

```

The predictor variable year has a strong left skew, and is unimodal. There are a significant number of outliers on the left. The median release year is 2007, and the IQR is 20.

```{r}
# Frequency Table
movies |>
  count(multi_language) |>
  mutate(proportion = n / sum(n)) |>
  kable(digits = 3)

# Bar Plot
movies |>
  ggplot(aes(x = factor(multi_language))) +  # Treat as categorical
  geom_bar() +
  labs(
    title = "Bar Plot of Multi-Language Movies",
    x = "Multi-Language (0 = No, 1 = Yes)", 
    y = "Count"
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes"))  # Adjust labels to be descriptive



```

About 70 percent of the data is in just English, and 30 percent features other languages.

```{r}
# Frequency Table
movies |>
  count(multi_genre) |>
  mutate(proportion = n / sum(n)) |>
  kable(digits = 3)

# Bar Plot
movies |>
  ggplot(aes(x = factor(multi_genre))) +  # Treat as categorical
  geom_bar() +
  labs(
    title = "Bar Plot of Multi-Genre Movies",
    x = "Multi-Genre (0 = No, 1 = Yes)", 
    y = "Count"
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes"))  # Adjust labels to be descriptive

```

About 84 percent of the movies are classified as more than one genre.

## Bivariate EDA

```{r}
movies |>
  select(vote_average, multi_language, budget, runtime, release_date, multi_genre) |>
  ggpairs()
```

## Carlie Section

```{r}
# spoken languages

movies |>
  ggplot(aes(x = multi_language, y = log_revenue)) +
  geom_boxplot() + 
  labs(x = "Number of Languages Spoken in Movie", 
       y = "log(Revenue)", 
       title = "Spoken Langauges in Movie vs. log(Revenue)"
       ) + 
  scale_x_discrete(labels = c("FALSE" = "One", "TRUE" = "More than One"))

```

```{r}
# genres 

movies |>
  ggplot(aes(x = multi_genre, y = log_revenue)) +
  geom_boxplot() + 
  labs(x = "Number of Genres in Movie", 
       y = "log(Revenue)", 
       title = "Genres in Movie vs. log(Revenue)"
       ) + 
  scale_x_discrete(labels = c("FALSE" = "One", "TRUE" = "More than One"))

```

```{r}
# budget

movies |>
  ggplot(aes(x = budget, y = log_revenue)) +
  geom_point() + 
  labs(
    title = "Budget vs. log(Revenue)",
    x = "Budget (in millions of dollars)",
    y = "log(Revenue)"
  ) +
  scale_x_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale()))
```

## Chrisie Section

```{r}
# vote
movies |>
  ggplot(aes(x = vote_average, y = log_revenue)) +
  geom_point() + 
  labs(
    title = "Vote Average vs. log(Revenue)",
    x = "Vote Average",
    y = "log(Revenue) (in billions of dollars)"
  ) +
  scale_y_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale()))
```

```{r}
# runtime
movies |>
  ggplot(aes(x = runtime, y = log_revenue)) +
  geom_point() + 
  labs(
    title = "Runtime vs. log(Revenue)",
    x = "Runtime (minutes)",
    y = "log(Revenue) (in billions of dollars)"
  ) +
  scale_y_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale()))
```

```{r}
# release date
movies |>
  ggplot(aes(x = year, y = log_revenue)) +
  geom_point() + 
  labs(
    title = "Release Date vs. log(Revenue)",
    x = "Release Date (year)",
    y = "log(Revenue) (in billions of dollars)"
  ) +
  scale_y_continuous(labels = scales::label_dollar(scale_cut = scales::cut_short_scale()))
```

## Interaction

::: callout-important
Before you submit, make sure your code chunks are turned off with `echo: false` and there are no warnings or messages with `warning: false` and `message: false` in the YAML.
:::
